systemfit: A Package for Estimating\\ Systems of Simultaneous Equations in \proglang{R}

### 2007-06-30 paper/trunk/JSS/response_A.tex

The method \code{lrtest.systemfit} is a wrapper function
to \code{lrtest.default} of the \pkg{lmtest} package now.
Its main task is to specify the names for the compared objects.

%
\subsection{Organize Internals of systemfit}
%   o organize internals of systemfit
% 
%     The internals of a fitted "systemfit" object could be re-organized. Currently,
%     there are a lot of single letter elements - more standard naming as in "lm"
%     objects etc. would be more desirable.
We have renamed many elements of a fitted \code{systemfit} object
to make the naming more standard.
A comparison with the elements returned by \code{lm}
is available in the paper (appendix~A).

@@

%     The names of various
%     arguments should be re-thought, e.g. "tx", "R.restr" and "q.rests" seem rather
%     arbitrary. Similar comments hold for other arguments, see also below for some
%     more comments.
The names of several arguments have been improved.
The arguments to impose parameter restrictions by $R \beta = q$
(former \code{R.restr} and \code{q.restr})
have been renamed to \code{restrict.matrix} and \code{restrict.rhs},
respectively
(following the arguments \code{hypothesis.matrix} and \code{rhs}
of the generic function \code{linear.hypothesis}).
The argument to impose parameter restrictions
by modifying (post-multiplying) the regressor matrix
(former \code{TX}) has been renamed to \code{restrict.regMat}.
Finally, the arguments to specify the methods
for the 3SLS estimation and to calculated the residual covariance matrix
(former \code{formula3sls} and \code{rcovformula})
have been renamed to \code{method3sls} and \code{methodRCov},
respectively.

%     As noted above, the standard in R seems to supply the formula(s)
%     as the first argument and to have then some `method' argument.
We have reversed the order of arguments \code{method} and \code{eqns}
of \code{systemfit()}.
It is no longer necessary to reverse the order of the arguments
of \code{systemfitClassic()},
because this function has been removed.

% 
%     Some more standard elements of a model fit should also be kept, in particular
%     the call so that update() can work with "systemfit" objects, e.g.,
%       fitsur <- systemfit("SUR", list(demand = eqDemand, supply = eqSupply))
%       fitsur$call <- call("systemfit", "SUR", list(demand = eqDemand, supply = eqSupply))
%       fitsurRmat <- update(fm, R.restr = Rmat, q.restr = qvec)
%     The call can easily be captured via match.call().
The call of \code{systemfit} is stored in the returned object now.
This allows the application of \code{update()} to \code{systemfit} objects.

%     Similarly, one could think 
%     about how the terms and formula can be preserved and possibly made available
%     in terms() and formula() calls. Other standard methods would also be nice to have,
%     e.g., df.residual() or logLik().
Following the suggestions of this referee,
the °model terms° (including the formulas) of each equation are preserved now.
We have added methods \code{terms.systemfit},
\code{terms}\hspace{0pt}\code{.systemfit}\hspace{0pt}\code{.equation},
\code{formula.systemfit}, and \code{formula.systemfit.equation}
to extract the °model terms° and formulas
of the system and of a single equation.
The new method \code{logLik} computes the log-likelihood value
of a fitted \code{systemfit} object.
Furthermore, the default method of \code{df.residual}
can extract the degrees of freedom of the residuals,
both of the system and of a single equation now.

%     (By the way: The row/column naming in fitted()
%     and residuals() should match, probably re-using equation and observation names.)
The data frames returned by \code{residuals.systemfit} and
\code{fitted.systemfit} have the same row names and column names now.

% 
%     The coefficient naming should be improved, e.g., "<eqname>_<varname>"
%     where <eqname> can be set by default to "eq1", "eq2" etc. but should use the
%     equation names (if supplied). Spaces in names should be avoided.
%     The coefficient naming in the systemfitClassic() interface could also be 
%     improved, see e.g. the output on p.27, l.546: "eq 2 capital.Chrysler" redundantly
%     codes the firm (both by eq and name) and "Chrysler_capital" or "Chryler.capital"
%     would be preferable.
The naming of the coefficients has been improved
according to the suggestions of this referee:
they are now \code{<eqname>}\verb!_!\code{<varname>},
where \code{<eqname>} is the equation name,
which is set to \code{"eq1"}, \code{"eq2"} etc.\
if no equation names are supplied by the user.
In the case of a (classical) °Seemingly Unrelated Regression° analysis
with panel-like data,
all redundant parts of the coefficient names have been removed:
the coefficient names in the vector of all coefficients are
\code{<individual>}\verb!_!\code{<varname>}
and the coefficient names in the vector of the coefficients of a single
individual are just \code{<varname>}.

### 2007-07-01 paper/trunk/JSS/response_A.tex

%     (By the way: The row/column naming in fitted()
%     and residuals() should match, probably re-using equation and observation names.)
The data frames returned by \code{residuals.systemfit} and
\code{fitted.systemfit} have the same row names and column names now.

% 
%     The coefficient naming should be improved, e.g., "<eqname>_<varname>"
%     where <eqname> can be set by default to "eq1", "eq2" etc. but should use the
%     equation names (if supplied). Spaces in names should be avoided.
%     The coefficient naming in the systemfitClassic() interface could also be 
%     improved, see e.g. the output on p.27, l.546: "eq 2 capital.Chrysler" redundantly
%     codes the firm (both by eq and name) and "Chrysler_capital" or "Chryler.capital"
%     would be preferable.
The naming of the coefficients has been improved
according to the suggestions of this referee:
they are now \code{<eqname>}\verb!_!\code{<varname>},
where \code{<eqname>} is the equation name,
which is set to \code{"eq1"}, \code{"eq2"} etc.\
if no equation names are supplied by the user.
In the case of a (classical) °Seemingly Unrelated Regression° analysis
with panel-like data,
all redundant parts of the coefficient names have been removed:
the coefficient names in the vector of all coefficients are
\code{<individual>}\verb!_!\code{<varname>}
and the coefficient names in the vector of the coefficients of a single
individual are just \code{<varname>}.

% 
%     It would also be nice to have some more control about the amount of output of
%     the summary() method. Currently, the output can become very long easily when the
%     number of equations increases. One could want to control the equations and
%     correlations used, e.g., by something like
%       summary(sfobject, equation = ..., cor = ...)
The user has more control about the amount of output of
\code{summary()} now.
We have added arguments \code{residCov} and \code{equations}
to function \code{print.summary.systemfit()}.
Argument \code{residCov} indicates
whether the residual correlation matrix,
the residual covariance matrix, and its determinant are printed;
argument \code{equations} indicates
whether summary results of each equation are printed
(or if just the coefficients are printed).
As users often use \code{summary()} without \code{print()},
we have added arguments \code{printResidCov} and \code{printEquations}
to \code{summary()},
which set default values for arguments \code{residCov} and \code{equations}
of function \code{print.summary.systemfit()}.

% 
%
\subsection{Computations in systemfit}
%   o computations in systemfit()
% 
%     A few computations within the systemfit() functions can be further streamlined.
%     A simple modification is to use
%       method %in% c("OLS", "WLS", ...)
%     instead of
%       method == "OLS" | method == "WLS" | method == ...
%     Furthermore, various marix computations can be made even more stable. Currently,
%     the code still has    
%       solve(crossprod(X), crossprod(X,Y))    
%       solve(crossprod(X))
%     which looks very similar to the code presented in Bates (2004). However, Bates
%     (2004) uses the "Matrix" package, which has specialized solve() for certain
%     matrix types. To perform this `by hand' without Matrix, one way would be re-using
%     lm.fit() for the first command. Alternatively, the QR or Cholesky decomposition
%     could be carried out using qr() or chol(). The latter command could be replaced
%     by
%      chol2inv(qr.R(qr(X)))
%      chol2inv(chol(crossprod(X)))
%     where chol() should be faster but qr() might be more accurate.
%
\subsection{Replication of Textbook Results (5)}
%   o Replication of textbook results (5)
% 
%     Having a section like this is excellent and should help teachers and students
%     for using systemfit in their lectures. Furthermore, it encourages reproducible
%     research which always deserves support. However, I would suggest to streamline
%     the section to make it more readable. The section should discuss its aims (where
%     I would add replication to reliability) and summarize the most important results
%     and experiences from the replication. Then, I would pick a `typical' example and
%     discuss it in some more detail with full R output. For the remaining examples,
%     the input suffices as long as there are no deviations of the systemfit results
%     from the published numbers.
% 
%     In addition to Table 14.1 and 14.2, I tried to replicate Table 14.3 with
%     ML estimates (using iterated FGLS) in Greene (2003) via 
%       systemfitClassic("WSUR", invest ~ value + capital, "firm", "year",
%                        data = GrunfeldGreene, rcovformula = 0, pooled = TRUE, maxiter = 1000)    
%     for the pooled model which yielded rather similar results; and via
%       systemfitClassic("SUR", invest ~ value + capital, "firm", "year",
%                        data = GrunfeldGreene, rcovformula = 0, maxiter = 1000)    
%     for the full model which is not unsimilar to Greene's results but also not very
%     close either. Is this a problem of systemfit or of Greene or of me fitting the
%     wrong model?
% 
%     Having two Grunfeld data sets is confusing as the data used by Theil is simply
%     a subset of the data used by Greene:
%       subset(GrunfeldGreene, firm %in% c("General Electric", "Westinghouse"))
%     Of course, there are different versions of this data set, but neither the paper
%     nor the manual pages of GrunfeldGreene and GrunfeldTheil appropriately discuss
%     the problem. From the manual and the paper, the reader is only able to feel
%     vaguely insecure about the appropriateness of the selected subsets. Also note
%     that a more complete (and slightly different) version of the data set is
%     available in package "Ecdat".
% 
%
\section{Minor Points and Typos}
We have corrected the paper according to all of these comments
except for °p.3, l.45° and the second part of °p.3, l.52°.

### 2007-07-01 paper/trunk/JSS/response_A.tex

% 
%     It would also be nice to have some more control about the amount of output of
%     the summary() method. Currently, the output can become very long easily when the
%     number of equations increases. One could want to control the equations and
%     correlations used, e.g., by something like
%       summary(sfobject, equation = ..., cor = ...)
The user has more control about the amount of output of
\code{summary()} now.
We have added arguments \code{residCov} and \code{equations}
to function \code{print.summary.systemfit()}.
Argument \code{residCov} indicates
whether the residual correlation matrix,
the residual covariance matrix, and its determinant are printed;
argument \code{equations} indicates
whether summary results of each equation are printed
(or if just the coefficients are printed).
As users often use \code{summary()} without \code{print()},
we have added arguments \code{printResidCov} and \code{printEquations}
to \code{summary()},
which set default values for arguments \code{residCov} and \code{equations}
of function \code{print.summary.systemfit()}.

% 
%
\subsection{Computations in systemfit}
%   o computations in systemfit()
% 
%     A few computations within the systemfit() functions can be further streamlined.
%     A simple modification is to use
%       method %in% c("OLS", "WLS", ...)
%     instead of
%       method == "OLS" | method == "WLS" | method == ...
We have streamlined the source code of the \pkg{systemfit} package.
For instance, we have followed the suggestion of this referee
and replaced all
°\code{method == "OLS" | method == "WLS" | method == ...}°
by °\code{method \%in\% c("OLS", "WLS", ...)}°.

%     Furthermore, various marix computations can be made even more stable. Currently,
%     the code still has    
%       solve(crossprod(X), crossprod(X,Y))    
%       solve(crossprod(X))
%     which looks very similar to the code presented in Bates (2004). However, Bates
%     (2004) uses the "Matrix" package, which has specialized solve() for certain
%     matrix types. To perform this `by hand' without Matrix, one way would be re-using
%     lm.fit() for the first command. Alternatively, the QR or Cholesky decomposition
%     could be carried out using qr() or chol(). The latter command could be replaced
%     by
%      chol2inv(qr.R(qr(X)))
%      chol2inv(chol(crossprod(X)))
%     where chol() should be faster but qr() might be more accurate.
%
\subsection{Replication of Textbook Results (5)}
%   o Replication of textbook results (5)
% 
%     Having a section like this is excellent and should help teachers and students
%     for using systemfit in their lectures. Furthermore, it encourages reproducible
%     research which always deserves support. However, I would suggest to streamline
%     the section to make it more readable. The section should discuss its aims (where
%     I would add replication to reliability) and summarize the most important results
%     and experiences from the replication. Then, I would pick a `typical' example and
%     discuss it in some more detail with full R output. For the remaining examples,
%     the input suffices as long as there are no deviations of the systemfit results
%     from the published numbers.
% 
%     In addition to Table 14.1 and 14.2, I tried to replicate Table 14.3 with
%     ML estimates (using iterated FGLS) in Greene (2003) via 
%       systemfitClassic("WSUR", invest ~ value + capital, "firm", "year",
%                        data = GrunfeldGreene, rcovformula = 0, pooled = TRUE, maxiter = 1000)    
%     for the pooled model which yielded rather similar results; and via
%       systemfitClassic("SUR", invest ~ value + capital, "firm", "year",
%                        data = GrunfeldGreene, rcovformula = 0, maxiter = 1000)    
%     for the full model which is not unsimilar to Greene's results but also not very
%     close either. Is this a problem of systemfit or of Greene or of me fitting the
%     wrong model?
% 
%     Having two Grunfeld data sets is confusing as the data used by Theil is simply
%     a subset of the data used by Greene:
%       subset(GrunfeldGreene, firm %in% c("General Electric", "Westinghouse"))
%     Of course, there are different versions of this data set, but neither the paper
%     nor the manual pages of GrunfeldGreene and GrunfeldTheil appropriately discuss
%     the problem. From the manual and the paper, the reader is only able to feel
%     vaguely insecure about the appropriateness of the selected subsets. Also note
%     that a more complete (and slightly different) version of the data set is
%     available in package "Ecdat".
The data set \code{GrunfeldTheil} has been removed,
because it is a subset of the data set \code{GrunfeldGreene}.
We have improved the discussion of different versions of Grunfeld's data set
in the documentation of \code{GrunfeldGreene}
and we have added this discussion in the paper.
Furthermore, we have added a note
that a correct version of this data set with five additional firms
is available as data set \code{Grunfeld} in the \pkg{Ecdat} package.

% 
%
\section{Minor Points and Typos}
We have corrected the paper according to all of these comments
except for °p.3, l.45° and the second part of °p.3, l.52°.

### 2007-07-02 paper/trunk/systemfit_estimation.tex
replaced all \widehat{\beta} by the macro \bHat, which is defined as \hat{\beta}
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@719 591db2bd-e0a0-4d22-8c19-2192dbb7da31
replaced all \widehat{X} by the macro \XHat
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@721 591db2bd-e0a0-4d22-8c19-2192dbb7da31
made naming of (estimated) residuals and (true) disturbances more consistent
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@779 591db2bd-e0a0-4d22-8c19-2192dbb7da31

The Ordinary Least Squares (OLS) estimator of the system 
is obtained by
\begin{equation}
   \widehat{\beta}_{OLS} = \left( X'X \right)^{-1} X'y .
   \label{eq:ols}
\end{equation}
These estimates are efficient only if the disturbance terms are not
contemporaneously correlated, which means
$\sigma_{ij} = 0 \; \forall \; i \neq j$.
If the whole system is treated as one single equation,
an estimator of the covariance matrix of the estimated parameters is
\begin{equation}
   \COVHat \left[ \widehat{\beta}_{OLS} \right] = \sHat^2 \left( X'X \right)^{-1} ,
   \label{eq:olsCovSameSigma}
\end{equation}
where $\sHat^2$ is the estimated variance of the residuals.
This assumes that the disturbances of all equations have the
same variance $(\E [ u_{it}^2 ] = \sigma^2 \, \forall \, i)$.

If the disturbance terms of the individual equations
are allowed to have different variances,
an estimator of the covariance matrix of the estimated parameters is
\begin{equation}
   \COVHat \left[ \widehat{\beta}_{OLS} \right] = \left( X' \OHat^{-1} X \right)^{-1} ,
   \label{eq:olsCovSingleSigma}
\end{equation}
where $\OHat = \SHat \otimes I$,
$\sHat_{ij} = 0 \; \forall \; i \neq j$ and
$\sHat_{ii} = \sHat_i^2$ is the estimated variance
of the disturbence term in the $i$th equation.

@@

The Weighted Least Squares (WLS) estimator allows for different variances
of the disturbance terms in the different equations
$( \sigma_i^2 \neq \sigma_j^2 \, \forall \, i \neq j$,
where $\sigma_i^2 = \E [ u_{it}^2 ])$.
The WLS estimator can be obtained by
\begin{equation}
   \widehat{\beta}_{WLS} = \left( X' \OHat^{-1} X \right)^{-1} X' \OHat^{-1} y ,
\end{equation}
where $\OHat = \SHat \otimes I$,
$\sHat_{ij} = 0 \; \forall \; i \neq j$ and
$\sHat_{ii} = \sHat_i^2$ is the estimated variance
of the disturbance term in the $i$th equation.%
\footnote{%
Theoretically, one should use the variances of the true residuals.
However, they are not known in most empirical applications.
Therefore, true variances are generally replaced by estimated variances
that are calculated from the residuals of a first-step OLS estimation.
}
Like the OLS estimates, these estimates are only efficient
if the disturbance terms are not contemporaneously correlated.
An estimator of the covariance matrix of the estimated parameters is
\begin{equation}
   \COVHat \left[ \widehat{\beta}_{WLS} \right] = \left( X' \OHat^{-1} X \right)^{-1} .
\end{equation}
If no cross-equation parameter restrictions are imposed,
the parameter estimates are equal to the OLS estimates.

@@

The SUR estimator can be obtained by
\begin{equation}
   \widehat{\beta}_{SUR} = \left( X' \OHat^{-1} X \right)^{-1} X' \OHat^{-1} y ,
   \label{eq:sur}
\end{equation}
where $\OHat = \SHat \otimes I$ and
$\SHat$ is the estimated covariance matrix of the disturbance terms.
An estimator of the covariance matrix of the estimated parameters is
\begin{equation}
   \COVHat \left[ \widehat{\beta}_{SUR} \right] = \left( X' \OHat^{-1} X \right)^{-1} .
   \label{eq:surCov}
\end{equation}

@@

At the first stage new ('fitted') regressors are obtained by
\begin{equation}
   \widehat{X_i} = H_i \left( H_i' H_i \right)^{-1} H_i' X .
\end{equation}
At the second stage, the unbiased two-stage least squares estimates
of $\beta$ are obtained by
\begin{equation}
   \widehat{\beta}_{2SLS} = \left( \widehat{X}' \widehat{X} \right)^{-1} 
   \widehat{X}' y .
   \label{eq:beta2sls}
\end{equation}
If the whole system is treated as one single equation,
an estimator of the covariance matrix of the estimated parameters is
\begin{equation}
   \COVHat \left[ \widehat{\beta}_{2SLS} \right] = \sHat^2 \left( \widehat{X}'
   \widehat{X} \right)^{-1} ,
   \label{eq:2slsCovSameSigma}
\end{equation}
where $\sHat^2$ is the estimated variance of the disturbance terms.
If the disturbance terms of the individual equations
are allowed to have different variances, 
an estimator of the covariance matrix of the estimated parameters is
\begin{equation}
   \COVHat \left[ \widehat{\beta}_{2SLS} \right] = \left( \widehat{X}' \OHat^{-1}
   \widehat{X} \right)^{-1} ,
   \label{eq:2slsCovSingleSigma}
\end{equation}
where $\OHat = \SHat \otimes I$,
$\sHat_{ij} = 0 \; \forall \; i \neq j$ and
$\sHat_{ii} = \sHat_i^2$ is the estimated variance
of the disturbance term in the $i$th equation.

@@

The Weighted Two-Stage Least Squares (W2SLS) estimator of the system 
is obtained by
\begin{equation}
   \widehat{\beta}_{W2SLS} = \left( \widehat{X}' \OHat^{-1} \widehat{X}
   \right)^{-1} \widehat{X}' \OHat^{-1} y ,
\end{equation}
where $\OHat = \SHat \otimes I$,
$\sHat_{ij} = 0 \; \forall \; i \neq j$ and
$\sHat_{ii} = \sHat_i^2$ is the estimated variance
of the disturbance term in the $i$th equation.
An estimator of the covariance matrix of the estimated parameters is
\begin{equation}
   \COVHat \left[ \widehat{\beta}_{W2SLS} \right] = \left( \widehat{X}' \OHat^{-1}
   \widehat{X} \right)^{-1} .
\end{equation}

@@

The standard 3SLS estimator can be obtained by
\begin{equation}
   \widehat{\beta}_{3SLS} = \left( \widehat{X}' \OHat^{-1} \widehat{X}
   \right)^{-1} \widehat{X}' \OHat^{-1} y ,
   \label{eq:3slsGls}
\end{equation}
where $\OHat = \SHat \otimes I$ and
$\SHat$ is the estimated covariance matrix of the disturbance terms.
An estimator of the covariance matrix of the estimated parameters is
\begin{equation}
   \COVHat \left[ \widehat{\beta}_{3SLS} \right] = \left( \widehat{X}' \OHat^{-1}
   \widehat{X} \right)^{-1} .
   \label{eq:cov3sls}
\end{equation}
While an unbiased 2SLS or W2SLS estimation requires only that
the instrumental variables and the disturbance terms of each single 
equation are uncorrelated
$( \E \left[ u_i' H_i \right]) = 0 \; \forall \; i )$,
\cite{schmidt90} points out that this estimator is only consistent 
if all disturbance terms and all instrumental variables are uncorrelated
$( \E \left[ u_i' H_j \right]) = 0 \; \forall \; i, j )$
with
\begin{equation}
   H =
   \left[ \begin{array}{cccc}
      H_1 & 0 & \cdots & 0\\
      0 & H_2 & \cdots & 0\\
      \vdots & \vdots & \ddots & \vdots\\
      0 & 0 & \cdots & H_G
   \end{array}\right] .
\end{equation}
Since there might be occasions where this cannot be avoided,
\cite{schmidt90} analyses other approaches to obtain 3SLS estimators.
One of these approaches based on instrumental variable estimation
(3SLS-IV) is
\begin{equation}
   \widehat{\beta}_{3SLS-IV} = \left( \widehat{X}' \OHat^{-1} X
   \right)^{-1} \widehat{X}' \OHat^{-1} y .
   \label{eq:3slsIv}
\end{equation}
An estimator of the covariance matrix of the estimated 3SLS-IV parameters is
\begin{equation}
   \COVHat \left[ \widehat{\beta}_{3SLS-IV} \right] = \left( \widehat{X}' \OHat^{-1}
   X \right)^{-1} .
\end{equation}
Another approach based on the Generalized Method of Moments (GMM)
estimator (3SLS-GMM) is
\begin{equation}
   \widehat{\beta}_{3SLS-GMM} = \left( X' H \left( H' \OHat H \right)^{-1}
   H' X \right)^{-1} X' H \left( H' \OHat H \right)^{-1} H' y .
   \label{eq:3slsGmm}
\end{equation}
An estimator of the covariance matrix of the estimated 3SLS-GMM parameters is
\begin{equation}
   \COVHat \left[ \widehat{\beta}_{3SLS-GMM} \right] =
   \left( X' H \left( H' \OHat H \right)^{-1} H' X \right)^{-1} .
\end{equation}
A fourth approach developed by \cite{schmidt90} himself is
\begin{equation}
   \widehat{\beta}_{3SLS-Schmidt} = \left( \widehat{X}' \OHat^{-1} \widehat{X}
   \right)^{-1} \widehat{X}' \OHat^{-1} 
   H \left( H' H \right)^{-1} H' y .
   \label{eq:3slsSchmidt}
\end{equation}
An estimator of the covariance matrix of these estimated parameters is
\begin{equation}
   \COVHat \left[ \widehat{\beta}_{3SLS-Schmidt} \right] =
   \left( \widehat{X}' \OHat^{-1}  \widehat{X} \right)^{-1}
   \widehat{X}' \OHat^{-1} H \left( H' H \right)^{-1} H' \OHat
   H \left( H' H \right)^{-1} H' \OHat^{-1} \widehat{X}
   \left( \widehat{X}' \OHat^{-1}  \widehat{X} \right)^{-1} .
\end{equation}
The econometrics software \proglang{EViews} uses
\begin{equation}
   \widehat{\beta}_{3SLS-EViews} = \widehat{\beta}_{2SLS} + 
   \left( \widehat{X}' \OHat^{-1} \widehat{X} \right)^{-1}
   \widehat{X}' \OHat^{-1} \left( y - X \widehat{\beta}_{2SLS} \right) ,
   \label{eq:3slsEViews}
\end{equation}
where $\widehat{\beta}_{2SLS}$ is the two-stage least squares estimator
as defined by (\ref{eq:beta2sls}).
\proglang{EViews} uses the standard 3SLS formula (\ref{eq:cov3sls}) to
calculate an estimator of the covariance matrix of the estimated parameters.

>>>>

The Ordinary Least Squares (OLS) estimator of the system 
is obtained by
\begin{equation}
   \bHat_{OLS} = \left( X'X \right)^{-1} X'y .
   \label{eq:ols}
\end{equation}
These estimates are efficient only if the disturbance terms are not
contemporaneously correlated, which means
$\sigma_{ij} = 0 \; \forall \; i \neq j$.
If the whole system is treated as one single equation,
an estimator of the covariance matrix of the estimated parameters is
\begin{equation}
   \COVHat \left[ \bHat_{OLS} \right] = \sHat^2 \left( X'X \right)^{-1} ,
   \label{eq:olsCovSameSigma}
\end{equation}
where $\sHat^2$ is the estimated variance of the disturbances.
This assumes that the disturbances of all equations have the
same variance $(\E [ u_{it}^2 ] = \sigma^2 \, \forall \, i)$.

If the disturbance terms of the individual equations
are allowed to have different variances,
an estimator of the covariance matrix of the estimated parameters is
\begin{equation}
   \COVHat \left[ \bHat_{OLS} \right] = \left( X' \OHat^{-1} X \right)^{-1} ,
   \label{eq:olsCovSingleSigma}
\end{equation}
where $\OHat = \SHat \otimes I$,
$\sHat_{ij} = 0 \; \forall \; i \neq j$ and
$\sHat_{ii} = \sHat_i^2$ is the estimated variance
of the disturbance term in the $i$th equation.

@@

The Weighted Least Squares (WLS) estimator allows for different variances
of the disturbance terms in the different equations
$( \sigma_i^2 \neq \sigma_j^2 \, \forall \, i \neq j$,
where $\sigma_i^2 = \E [ u_{it}^2 ])$.
The WLS estimator can be obtained by
\begin{equation}
   \bHat_{WLS} = \left( X' \OHat^{-1} X \right)^{-1} X' \OHat^{-1} y ,
\end{equation}
where $\OHat = \SHat \otimes I$,
$\sHat_{ij} = 0 \; \forall \; i \neq j$ and
$\sHat_{ii} = \sHat_i^2$ is the variance
of the (estimated) residuals in the $i$th equation.
Theoretically, it would be better to use the variances of
the (true) disturbances $( \sigma_{ii} )$.
However, they are not known in most empirical applications.
Therefore, true variances are generally replaced by estimated variances
$( \sHat_{ii} )$
that are calculated from the residuals of a first-step OLS estimation
(see section~\ref{sec:residcov}).

Like the OLS estimates, these estimates are only efficient
if the disturbance terms are not contemporaneously correlated.
An estimator of the covariance matrix of the estimated parameters is
\begin{equation}
   \COVHat \left[ \bHat_{WLS} \right] = \left( X' \OHat^{-1} X \right)^{-1} .
\end{equation}
If no cross-equation parameter restrictions are imposed,
the parameter estimates are equal to the OLS estimates.

@@

The SUR estimator can be obtained by
\begin{equation}
   \bHat_{SUR} = \left( X' \OHat^{-1} X \right)^{-1} X' \OHat^{-1} y ,
   \label{eq:sur}
\end{equation}
where $\OHat = \SHat \otimes I$ and
$\SHat$ is the estimated covariance matrix of the disturbance terms.
An estimator of the covariance matrix of the estimated parameters is
\begin{equation}
   \COVHat \left[ \bHat_{SUR} \right] = \left( X' \OHat^{-1} X \right)^{-1} .
   \label{eq:surCov}
\end{equation}

@@

At the first stage new ('fitted') regressors are obtained by
\begin{equation}
   \XHat_i = H_i \left( H_i' H_i \right)^{-1} H_i' X .
\end{equation}
At the second stage, the unbiased two-stage least squares estimates
of $\beta$ are obtained by
\begin{equation}
   \bHat_{2SLS} = \left( \XHat' \XHat \right)^{-1}
   \XHat' y .
   \label{eq:beta2sls}
\end{equation}
If the whole system is treated as one single equation,
an estimator of the covariance matrix of the estimated parameters is
\begin{equation}
   \COVHat \left[ \bHat_{2SLS} \right] = \sHat^2 \left( \XHat'
   \XHat \right)^{-1} ,
   \label{eq:2slsCovSameSigma}
\end{equation}
where $\sHat^2$ is the estimated variance of the disturbance terms.
If the disturbance terms of the individual equations
are allowed to have different variances, 
an estimator of the covariance matrix of the estimated parameters is
\begin{equation}
   \COVHat \left[ \bHat_{2SLS} \right] = \left( \XHat' \OHat^{-1}
   \XHat \right)^{-1} ,
   \label{eq:2slsCovSingleSigma}
\end{equation}
where $\OHat = \SHat \otimes I$,
$\sHat_{ij} = 0 \; \forall \; i \neq j$ and
$\sHat_{ii} = \sHat_i^2$ is the estimated variance
of the disturbance term in the $i$th equation.

@@

The Weighted Two-Stage Least Squares (W2SLS) estimator of the system 
is obtained by
\begin{equation}
   \bHat_{W2SLS} = \left( \XHat' \OHat^{-1} \XHat
   \right)^{-1} \XHat' \OHat^{-1} y ,
\end{equation}
where $\OHat = \SHat \otimes I$,
$\sHat_{ij} = 0 \; \forall \; i \neq j$ and
$\sHat_{ii} = \sHat_i^2$ is the estimated variance
of the disturbance term in the $i$th equation.
An estimator of the covariance matrix of the estimated parameters is
\begin{equation}
   \COVHat \left[ \bHat_{W2SLS} \right] = \left( \XHat' \OHat^{-1}
   \XHat \right)^{-1} .
\end{equation}

@@

The standard 3SLS estimator can be obtained by
\begin{equation}
   \bHat_{3SLS} = \left( \XHat' \OHat^{-1} \XHat
   \right)^{-1} \XHat' \OHat^{-1} y ,
   \label{eq:3slsGls}
\end{equation}
where $\OHat = \SHat \otimes I$ and
$\SHat$ is the estimated covariance matrix of the disturbance terms.
An estimator of the covariance matrix of the estimated parameters is
\begin{equation}
   \COVHat \left[ \bHat_{3SLS} \right] = \left( \XHat' \OHat^{-1}
   \XHat \right)^{-1} .
   \label{eq:cov3sls}
\end{equation}
While an unbiased 2SLS or W2SLS estimation requires only that
the instrumental variables and the disturbance terms of each single 
equation are uncorrelated
$( \E \left[ u_i' H_i \right]) = 0 \; \forall \; i )$,
\cite{schmidt90} points out that this estimator is only consistent 
if all disturbance terms and all instrumental variables are uncorrelated
$( \E \left[ u_i' H_j \right]) = 0 \; \forall \; i, j )$
with
\begin{equation}
   H =
   \left[ \begin{array}{cccc}
      H_1 & 0 & \cdots & 0\\
      0 & H_2 & \cdots & 0\\
      \vdots & \vdots & \ddots & \vdots\\
      0 & 0 & \cdots & H_G
   \end{array}\right] .
\end{equation}
Since there might be occasions where this cannot be avoided,
\cite{schmidt90} analyses other approaches to obtain 3SLS estimators.
One of these approaches based on instrumental variable estimation
(3SLS-IV) is
\begin{equation}
   \bHat_{3SLS-IV} = \left( \XHat' \OHat^{-1} X
   \right)^{-1} \XHat' \OHat^{-1} y .
   \label{eq:3slsIv}
\end{equation}
An estimator of the covariance matrix of the estimated 3SLS-IV parameters is
\begin{equation}
   \COVHat \left[ \bHat_{3SLS-IV} \right] = \left( \XHat' \OHat^{-1}
   X \right)^{-1} .
\end{equation}
Another approach based on the Generalized Method of Moments (GMM)
estimator (3SLS-GMM) is
\begin{equation}
   \bHat_{3SLS-GMM} = \left( X' H \left( H' \OHat H \right)^{-1}
   H' X \right)^{-1} X' H \left( H' \OHat H \right)^{-1} H' y .
   \label{eq:3slsGmm}
\end{equation}
An estimator of the covariance matrix of the estimated 3SLS-GMM parameters is
\begin{equation}
   \COVHat \left[ \bHat_{3SLS-GMM} \right] =
   \left( X' H \left( H' \OHat H \right)^{-1} H' X \right)^{-1} .
\end{equation}
A fourth approach developed by \cite{schmidt90} himself is
\begin{equation}
   \bHat_{3SLS-Schmidt} = \left( \XHat' \OHat^{-1} \XHat
   \right)^{-1} \XHat' \OHat^{-1}
   H \left( H' H \right)^{-1} H' y .
   \label{eq:3slsSchmidt}
\end{equation}
An estimator of the covariance matrix of these estimated parameters is
\begin{equation}
   \COVHat \left[ \bHat_{3SLS-Schmidt} \right] =
   \left( \XHat' \OHat^{-1}  \XHat \right)^{-1}
   \XHat' \OHat^{-1} H \left( H' H \right)^{-1} H' \OHat
   H \left( H' H \right)^{-1} H' \OHat^{-1} \XHat
   \left( \XHat' \OHat^{-1}  \XHat \right)^{-1} .
\end{equation}
The econometrics software \proglang{EViews} uses
\begin{equation}
   \bHat_{3SLS-EViews} = \bHat_{2SLS} +
   \left( \XHat' \OHat^{-1} \XHat \right)^{-1}
   \XHat' \OHat^{-1} \left( y - X \bHat_{2SLS} \right) ,
   \label{eq:3slsEViews}
\end{equation}
where $\bHat_{2SLS}$ is the two-stage least squares estimator
as defined by (\ref{eq:beta2sls}).
\proglang{EViews} uses the standard 3SLS formula (\ref{eq:cov3sls}) to
calculate an estimator of the covariance matrix of the estimated parameters.
